---
title: "Activity 1: LA Housing Data"
author: "Alice Chang"
output: pdf_document
---

```{r setup, include=FALSE}
URL <- "http://andrewpbray.github.io/data/LA.csv"
LA <- read.csv(URL)
```



```{r}
m1 <- lm(log(price) ~ log(sqft) + bed + city, data = LA)
summary(m1)

m2 <- lm(log(price) ~ log(sqft) + bed + city + log(sqft):city, data = LA)
summary(m2)

m3 <- lm(log(price) ~ log(sqft) + bed + log(sqft):bed, data = LA)
summary(m3)

```
The geometry of the first model seems to be four parallel planes, while the geometry of the second model would be four planes with different slopes. 

Beverly Hills appears to be the reference level for city?

For every increase in bedrooms in a home -- holding all other variables constant -- the log price of the house decreases by 5.13. This makes sense because by holding the square footage of a house constant and adding more rooms into the limited space, it is likely that the value of the house will go down. 
## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
