---
title: "Lab 9"
author: "Alice Chang"
output: pdf_document
---


## The Letter "R""
```{r, echo = FALSE}
library(dplyr)
  
d <- read.csv("https://raw.githubusercontent.com/stat-learning/course-materials/master/data/handwritten.csv")


#Create new dataset for "r"
r_data <- d %>%
  filter(letter == "r")

#Plot a few images for "r"
plot_letter <- function(x, hasletter = TRUE) {
  if(hasletter) {
    a <- as.numeric(x[, -1])
  }else{a <- as.numeric(x)}
  m <- matrix(a, nrow = 8, byrow = TRUE)
  m <- t(apply(m, 2, rev)) # rotate matrix
  par(mar = rep(0, 4))
  image(m, axes = FALSE, col = rev(grey(seq(0, 1, length = 256)))) #this should be a divergent palette
  box()
}


plot_letter(r_data[1,], hasletter = TRUE)
plot_letter(r_data[2,], hasletter = TRUE)


#Compute and visualize mean image for letter
r_data <- r_data[, -1]

mean_r <- colMeans(r_data)
mean_r


plot_letter(mean_r, hasletter = FALSE)

#Perform PCA
pca1 <- prcomp(r_data, scale = TRUE)
names(pca1)

#Create biplot
phi <- pca1$rotation
phi
biplot(pca1)

#Plot first several PCs, learn about PC1 AND PC2
plot_letter(phi[,1], hasletter = FALSE)
plot_letter(phi[,2], hasletter = FALSE)



#Plot the PCs
d_2<- as.data.frame(pca1$x)
library(ggplot2)
p1 <- ggplot(d_2, aes(x = PC1, y = PC2)) + geom_point(alpha = .4) + theme_bw()
p1

#Construct Scree Plor
frame <- data.frame(PC = 1:20,
                PVE = pca1$sdev[1:20]^2 /
                  sum(pca1$sdev[1:20]^2))
ggplot(frame, aes(x = PC, y = PVE)) +
  geom_line() + 
  geom_point()

#Scatterplot of selected observations

pc_grid <- function(pca, data) {
  d <- data
  grid_points <- as.matrix(expand.grid(seq(-1.5, 1.5, length.out = 5), 
                                       seq(-1.5, 1.5, length.out = 5)))
  pc_points <- pca$x[, 1:2]
  nearest_ind <- rep(NA, nrow(grid_points))
  for(i in 1:nrow(grid_points)) {
    gp <- matrix(rep(grid_points[i, ], nrow(pc_points)), 
                 ncol = 2, byrow = TRUE)
    nearest_ind[i] <- which.min(rowSums((pc_points - gp)^2))
  }
  nearest_grid <- data.frame(d[nearest_ind, ])
  par(mfrow = c(5, 5))
  regrid <- c(21:25, 16:20, 11:15, 6:10, 1:5)
  for(i in regrid) {
    plot_letter(nearest_grid[i, ])
  }
}

pc_grid(pca1, r_data)

#Image construction 
z <- pca1$x
       #1 by m
z_tilda <- z[ ,1:10]

        #64 by ten
phi <- pca1$rotation
phi_tilda <- phi[,1:10]
image1 <- t(mean_r) + z_tilda [1, ] %*% t(phi_tilda)
image2 <- t(mean_r) + z_tilda [2, ] %*% t(phi_tilda)
plot_letter(image1, hasletter = FALSE)
plot_letter(image2, hasletter = FALSE)

#Fuller image construction
z_fuller <- z[ ,1:60]
phi_fuller <- phi[,1:60]

full_image1 <- t(mean_r) + z_fuller[1,] %*% t(phi_fuller)
plot_letter(full_image1, hasletter = FALSE)

full_image2 <- t(mean_r) + z_fuller[2,] %*% t(phi_fuller)
plot_letter(full_image2, hasletter = FALSE)



```
In this dataset, the rows seem to represent the letter classification of each image, and the columns represent each of the pixels in the image. 

Exploring the letter grid, PC1 (horizontal axis) seems to be encoding for how left-aligned the letters are (or how much the letters ascend upwards diagonally from left to right) in the image. PC2(vertical axis) could be encoding for the width of the image that is filled by the letter. 

The fuller image construction, yields the more accurate letter result. 


## The Letter "L"
```{r}
l_data <- d %>%
  filter(letter == "l")


#Compute/visualize mean image for "l"
l_data <- l_data[, -1]
mean_l <- colMeans(l_data)
plot_letter(mean_l, hasletter = FALSE)

#Perform PCA
pca2 <- prcomp(l_data, scale = TRUE)

#Construct Scree Plot
frame_l <- data.frame(PC = 1:20,
                PVE = pca1$sdev[1:20]^2 /
                  sum(pca1$sdev[1:20]^2))
ggplot(frame_l, aes(x = PC, y = PVE)) +
  geom_line() + 
  geom_point()

#Plot first several PCs
phi_l <- pca2$rotation
plot_letter(phi_l[,1], hasletter = FALSE)
plot_letter(phi_l[,2], hasletter = FALSE)

#Plot the PCs
d_3<- as.data.frame(pca2$x)
library(ggplot2)
p2 <- ggplot(d_3, aes(x = PC1, y = PC2)) + geom_point(alpha = .4) + theme_bw()
p2

#Plot letter grid
pc_grid(pca2, l_data)

#Image construction
z_l <- pca2$x
z_l_tilda <- z_l[,1:10]
phi_l_tilda <- phi_l[,1:10]
image1_l <- t(mean_l) + z_l_tilda [1, ] %*% t(phi_l_tilda)
image2_l <- t(mean_l) + z_l_tilda [2, ] %*% t(phi_l_tilda)
plot_letter(image1_l, hasletter = FALSE)
plot_letter(image2_l, hasletter = FALSE)

#Fuller image construction
z_l_full <- z_l[,1:60]
phi_l_full <- phi_l[,1:60]
image1_l_full <- t(mean_l) + z_l_full [1, ] %*% t(phi_l_full)
image2_l_full <- t(mean_l) + z_l_full [2, ] %*% t(phi_l_full)
plot_letter(image1_l_full, hasletter = FALSE)
plot_letter(image2_l_full, hasletter = FALSE)




```
Examining the letter grid, PC1 (horizontal axis) seems to be encoding for how centered the letter is in the image. PC2(vertical axis) could be encoding for the straightness of the letter. As observed previously, a fuller image contstruction yields a more accurate result.